The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-03-23 src/ms.tex
Lim edits for Sections 2.3 and 2.4

An important focus of \astropy has been to increase the precision and accuracy
of its functionality, while, where appropriate, making sure results obtained
using older versions of \astropypkg are reproducible. Within each section below
we include in our descriptions how old results may be reproduced. The primary
means are with options in configuration files and settings on runtime configuration
objects called \astropyScienceState. Notable improvements to \astropypkg have
been made to \astropytime, \astropyconstants, \astropycoordinates, and
\astropycosmology, and we describe each in turn.

\paragraph{Time}
In astronomy, time accuracy down to the (nano)second is frequently important.
Whether for planning observations or crunching pulsar or VLBI data, missing
seconds meaningfully impact results. In past versions of \astropypkg, when using
\astropyTime\, leap seconds had to be manually applied; now, leap
seconds are applied automatically. Moreover, \astropypkg updates internal time
data files to ensure that the correct leap second adjustment is always used. For
reproducibility with older code, \astropyLeapSeconds\ may also be
manually applied.

@@

Measurements of physical constants (and the units defined from them) improve
over time. Periodically the standardized systems of units and constants are
updated to reflect these improvements. For instance, in 2019 the SI system was
redefined \citep{NIST2019}, with an accompanying update to the physical
constants in
\href{https://codata.org/initiatives/data-science-and-stewardship/fundamental-physical-constants/}{\texttt{SI/CODATA
2018}}. \astropyconstants now defaults to using the \texttt{SI/CODATA 2018}
values, with the units in \astropyunits based on these constants. Most of
\astropypkg and affiliate packages build upon \astropyunits, so this update
affects the entire \astropy ecosystem.

For reproducibility, \astropypkg allows the constants' (and therefore units')
definitions to be rolled back to prior values, e.g., to the \texttt{SI/CODATA
2014} values. For work sensitive to the values of the fundamental constants we
recommend including an \astropy configuration file with the work, which
specifies the set of constants used.

@@

Astropy provides a
\href{http://docs.astropy.org/en/stable/api/astropy.cosmology.default_cosmology.html}{configurable
default cosmology}, which is used in calculations done in a cosmological
context. The default \astropyCosmology\ has been updated to the Planck 2018
parameters. For reproducibility, this cosmology may be set to old defaults, such
as the WMAP Three Year values. The default cosmology is dynamically configurable
using \astropyScienceState, meaning a set of calculations may be run using
different assumed cosmologies, and results compared between the two.

@@

Being able to describe the position of objects, i.e., their coordinates, is
fundamental across many knowledge domains. \astropycoordinates allows one to
work with low-level positional and velocity data all the way to high-level
coordinate objects with reference frames, atmospheric information, and more. A
central feature of \astropycoordinates is the ability to transform data between
reference frames, e.g., \astropyICRS\ \citep{ICRS:1997} to \astropyGalacticLSR\
\citep{GalacticLSR:2010}. However, there have been some historical limitations
for spatially proximate transformations, impacting the usefulness of
\astropycoordinates for ground-based telescopes and astrometry within the
solar system. Transformations in the \astropyAltAz\ frame were reasonably
precise for very distant objects, but wrong by up to several arcseconds for
e.g., the location of the moon. Now these transformations are much more precise,
down to the milliarcsecond level. Similar precision improvements were made to
the
\astropyapi{coordinates.builtin_frames.HADec}{Hour Angle-Declination}
frame transformations. Additionally,
\astropyapi{coordinates.builtin_frames.BaseEclipticFrame}{Ecliptic frames}
and associated transformations have been updated to correctly reflect the ``true''
and ``mean" terminology.

@@

As features are added and corner cases dealt with, performance often suffers.
Hence, pull requests that improve performance are welcomed, and were even the
main goal for one release cycle. For most sub-packages, performance was
improved in the \python\ code, but for some, the most time-critical pieces were
rewritten in~C (e.g., for convolution of images, sigma-clipping, and converting
time strings to binary). Furthermore, a particular effort was made to ensure
\astropypkg\ is thread-safe, so that it can be used on supercomputer clusters.

>>>>

An important focus of \astropypkg has been to increase the numerical precision and accuracy
of its functionality, while, where appropriate, making sure results obtained
using older versions of \astropypkg are reproducible. Within each section below,
we include in our descriptions how old results may be reproduced. The primary
means are with options in configuration files and settings on runtime configuration
objects called \astropyScienceState. Notable improvements to \astropypkg have
been made to \astropytime, \astropyconstants, \astropycoordinates, and
\astropycosmology, and we describe each in turn.

\paragraph{Time}
In astronomy, time accuracy down to the (nano)second is frequently important.
Whether for planning observations or crunching pulsar or VLBI data, missing
seconds meaningfully impact results. In past versions of \astropypkg, when using
\astropyTime, leap seconds had to be manually applied; now, leap
seconds are applied automatically. Moreover, \astropypkg updates internal time
data files to ensure that the correct leap second adjustment is always used. For
reproducibility with older code, \astropyLeapSeconds\ may also be
manually applied.

@@

Measurements of physical constants (and the units defined from them) improve
over time. Periodically the standardized systems of units and constants are
updated to reflect these improvements. For instance, in 2019 the SI system was
redefined \citep{NIST2019}, with an accompanying update to the physical
constants in
\href{https://codata.org/initiatives/data-science-and-stewardship/fundamental-physical-constants/}{\texttt{SI/CODATA
2018}}. \astropyconstants now defaults to using the \texttt{SI/CODATA 2018}
values, with the units in \astropyunits based on these constants. Most of
\astropy and affiliated packages build upon \astropyunits, so this update
affects the entire \astropy ecosystem.

For reproducibility, \astropypkg allows the constants' (and therefore units')
definitions to be rolled back to prior values, e.g., to \texttt{SI/CODATA
2014}. For work sensitive to the values of the fundamental constants, we
recommend including an \astropypkg configuration file (i.e., \texttt{astropy.cfg})
with the work, which specifies the set of constants used.

@@

\astropypkg provides a
\href{http://docs.astropy.org/en/stable/api/astropy.cosmology.default_cosmology.html}{configurable
default cosmology}, which is used in calculations done in a cosmological
context. The default \astropyCosmology\ has been updated to the Planck 2018
parameters. For reproducibility, this cosmology may be set to old defaults, such
as the WMAP Three Year values. The default cosmology is dynamically configurable
using \astropyScienceState, meaning a set of calculations may be run using
different assumed cosmologies, and results compared between the two.

@@

Being able to describe the coordinates of objects is
fundamental across many knowledge domains. \astropycoordinates allows one to
work with low-level positional and velocity data all the way to high-level
coordinate objects with reference frames, atmospheric information, etc. A
central feature of \astropycoordinates is the ability to transform data between
reference frames, e.g., \astropyICRS\ \citep{ICRS:1997} to \astropyGalacticLSR\
\citep{GalacticLSR:2010}. However, there have been some historical limitations
for spatially proximate transformations, impacting the usefulness of
\astropycoordinates for ground-based telescopes and astrometry within the
solar system. Transformations in the \astropyAltAz\ frame were reasonably
precise for very distant objects, but wrong by up to several arcseconds for,
e.g., the location of the moon. Now these transformations are much more precise,
down to the milliarcsecond level. Similar precision improvements were made to
the
\astropyapi{coordinates.builtin_frames.HADec}{Hour Angle-Declination}
frame transformations. Additionally,
\astropyapi{coordinates.builtin_frames.BaseEclipticFrame}{Ecliptic frames}
and associated transformations have been updated to correctly reflect the ``true''
and ``mean" terminology.

@@

In some cases, feature enhancements and corner-case handling were added
at the cost of performance.
Hence, \github pull requests that improve performance are welcomed, and were even the
main goal for v3.1 release. For most sub-packages, performance was
improved in the \python\ code, but for some, the most time-critical pieces were
rewritten in~C (e.g., for convolution of images, sigma-clipping, and converting
time strings to binary). Furthermore, a particular effort was made to ensure
\astropypkg\ is thread-safe, so that it can be used on supercomputing clusters.
